Self Regulation Ontology Retest Data Report

### 2017-04-22 SRO_Report.Rmd
began refactoring code to make bootstrapping more efficient and report shorter and more compact

## Correlations for all cleaned meaningful variables
Spearman only

@@

cor_df <- data.frame(spearman = rep(NA, length(numeric_cols)))

@@

for(i in 1:length(numeric_cols)){
  cor_df[numeric_cols[i], 'spearman'] <- cor(retest_data[,numeric_cols[i]],retest_subs_test_data[,numeric_cols[i]], method = 'spearman', use = "pairwise.complete.obs")
}

@@

Mean Spearman correlation and the number of variables for both the surveys and the tasks
```{r}
datatable(cor_df %>%
  group_by(task) %>%
  summarise(mean_spearman = mean(spearman),
            num_vars = n()))
```

### Distributions of correlations by task vs. survey

Spearman correlations for surveys are noticably higher than task reliabilities.

@@

@@

#### Rank order of correlations

##### For tasks
```{r}
datatable(cor_df %>%
  filter(task == 'task') %>%
  select(dv, spearman) %>%
  arrange(-spearman))
```

##### For surveys

- Demographics are excluded from this list of variables (because they are saved for predictive analyses).  
- Holt and Laury test-retest is the worst for *surveys*. But note that this is not a survey that has been studied extensively psychometrically like others. This is more similar to our tasks (one that is used more often by economists). In previous data I have worked with retest reliablities for this task was ~.35 so the results here are in line with this.  
- Otherwise the lowest survey reliabilities are 0.5.
```{r}
datatable(cor_df %>%
  filter(task == 'survey') %>%
  select(dv, spearman) %>%
  arrange(-spearman))
```

## ICC for all cleaned meaningful variables

Based on [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) ICC(3,k) does not take in to account within subject differences between two time points (i.e. the fixed effect of time/systematic error). Thus, it is well approximated by Pearson's r and subject to similar criticisms. [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) suggests reporting at least this systematic error effect size if one chooses to report with ICC(3,k). Based on his conclusions here I report:  
- ICC(3,k): As Dave clarified this ranges from 1 to -1/(number of repeated measures -1) so in our case this range would be [-1, 1]; larger values would mean that the two scores of a subject for a given measure are more similar to each other than they are to scores of other people  
- partial $\eta^2$ for time ($SS_{time}/SS_{within}$): effect size of time   
- SEM ($\sqrt(MS_{error})$): standard error of measurement; the smaller the better

```{r}
match_t1_t2 <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', format = "long"){
  df = merge(t1_df[,c(merge_var, dv_var)], t2_df[,c(merge_var, dv_var)], by = merge_var)

get_icc <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', match=TRUE, matched_df){
  if(match==TRUE){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide')
  }
  else{
    df = matched_df
  }
  df = df %>% select(-dv, -sub_id)
  icc = ICC(df)
  icc_3k = icc$results['Average_fixed_raters', 'ICC']
  return(icc_3k)
  }

get_eta <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', match=TRUE, matched_df){
  if(match==TRUE){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var)
  }
  else{
    df = matched_df
  }
  mod = summary(aov(score~Error(sub_id)+time, df))
  ss_time = as.data.frame(unlist(mod$`Error: Within`))['Sum Sq1',]
  ss_error = as.data.frame(unlist(mod$`Error: Within`))['Sum Sq2',]
  eta = ss_time/(ss_time+ss_error)
  return(eta)
  }

get_sem <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', match=TRUE, matched_df){
  if(match==TRUE){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var)
  }
  else{
    df = matched_df
  }
  mod = summary(aov(score~Error(sub_id)+time, df))
  ms_error = as.data.frame(unlist(mod$`Error: Within`))['Mean Sq2',]
  sem = sqrt(ms_error)
  return(sem)
}
```

```{r}
#Calculate ICC df
icc_df <- data.frame(icc = rep(NA, length(numeric_cols)), 
                     eta_sq = rep(NA, length(numeric_cols)),
                     sem = rep(NA, length(numeric_cols)))

icc_df$dv = row.names(icc_df)
row.names(icc_df) = seq(1:nrow(icc_df))
icc_df$task = 'task'
icc_df[grep('survey', icc_df$dv), 'task'] = 'survey'
icc_df = icc_df %>%
  select(dv, task, icc, eta_sq, sem)
row.names(icc_df) = NULL
```

Median reliability and the number of variables for both the surveys and the tasks:
As seen with Spearman $\rho$'s as well the ICC's are higher for surveys.  
Systematic differences between the two measurements ($\eta^2$) are higher for tasks than surveys (potentially indicating learning effects?).  
Unexpectedly, median standard error of measurement is higher for surveys but the distribution below suggests that this is a side effect of having too many task measures with very low SEMs.
```{r}
datatable(icc_df %>%
  group_by(task) %>%
  summarise(median_icc = median(icc),
            median_eta_sq = median(eta_sq),
            median_sem = median(sem),
            num_vars = n()))
```

#### Rank order of ICCs, partial $\eta^2$ and SEM

Note that these tables can be sorted for any of the columns.

###### For tasks

```{r}
datatable(icc_df %>%
  filter(task == 'task') %>%
  left_join(cor_df[,c('spearman','dv')], by = 'dv') %>%
  arrange(-icc) %>%
  select(dv, -task, icc, eta_sq, sem, spearman))
```

###### For surveys

```{r}
datatable(icc_df %>%
  filter(task == 'survey') %>%
  left_join(cor_df[,c('spearman','dv')], by = 'dv') %>%
  arrange(-icc) %>%
  select(dv, -task, icc, eta_sq, sem, spearman))
```

### Distributions of ICC's by task vs. survey

@@

### Distributions of partial $\eta^2$ by task vs. survey

@@

### Distributions of SEM's by task vs. survey

@@

extract_sample_data = function(sample, var_name, df, worker_col = "sub_id"){
  return(df[df[,worker_col] %in% sample, var_name])
}

sample_workers()
extract_sample_data(df=retest_data)
extract_sample_data(df=retest_subs_test_data)
```

>>>>

## Reliability metrics 
For correlations we report Spearman's $\rho$'s only.  
Based on [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) ICC(3,k) does not take in to account within subject differences between two time points (i.e. the fixed effect of time/systematic error). Thus, it is well approximated by Pearson's r and subject to similar criticisms. [Weir (2005)](https://pdfs.semanticscholar.org/d99a/790cce43f7f20d742f9d379b79de4f767740.pdf) suggests reporting at least this systematic error effect size if one chooses to report with ICC(3,k). Based on his conclusions here I report:  
- ICC(3,k): As Dave clarified this ranges from 1 to -1/(number of repeated measures -1) so in our case this range would be [-1, 1]; larger values would mean that the two scores of a subject for a given measure are more similar to each other than they are to scores of other people  
- partial $\eta^2$ for time ($SS_{time}/SS_{within}$): effect size of time   
- SEM ($\sqrt(MS_{error})$): standard error of measurement; the smaller the better

@@

match_t1_t2 <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', format = "long", sample = 'full', sample_vec){

  if(sample == 'full'){
   df = merge(t1_df[,c(merge_var, dv_var)], t2_df[,c(merge_var, dv_var)], by = merge_var) 
  }
  else{
    df = merge(t1_df[t1_df[,merge_var] %in% sample_vec, c(merge_var, dv_var)], t2_df[t2_df[,merge_var] %in% sample_vec, c(merge_var, dv_var)],
    by=merge_var)
  }

get_spearman = function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){

  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide')
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide', sample='bootstrap', sample_vec = sample_vec)
  }

  rho = cor(df$`1`, df$`2`, method='spearman')

  return(rho)
}

get_icc <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){
  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide')
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, format='wide', sample='bootstrap', sample_vec = sample_vec)
  }

  df = df %>% select(-dv, -sub_id)
  icc = ICC(df)
  icc_3k = icc$results['Average_fixed_raters', 'ICC']
  return(icc_3k)
  }

get_eta <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){
  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var)
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, sample='bootstrap', sample_vec = sample_vec)
  }

  mod = summary(aov(score~Error(sub_id)+time, df))
  ss_time = as.data.frame(unlist(mod$`Error: Within`))['Sum Sq1',]
  ss_error = as.data.frame(unlist(mod$`Error: Within`))['Sum Sq2',]
  eta = ss_time/(ss_time+ss_error)
  return(eta)
  }

get_sem <- function(dv_var, t1_df = retest_subs_test_data, t2_df = retest_data, merge_var = 'sub_id', sample='full', sample_vec){
  if(sample=='full'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var)
  }
  else if(sample=='bootstrap'){
    df = match_t1_t2(dv_var, t1_df = t1_df, t2_df = t2_df, merge_var = merge_var, sample='bootstrap', sample_vec = sample_vec)
  }
  mod = summary(aov(score~Error(sub_id)+time, df))
  ms_error = as.data.frame(unlist(mod$`Error: Within`))['Mean Sq2',]
  sem = sqrt(ms_error)
  return(sem)
}
```

```{r}
cor_df <- data.frame(spearman = rep(NA, length(numeric_cols)))

@@

for(i in 1:length(numeric_cols)){
  cor_df[numeric_cols[i], 'spearman'] <- get_spearman(numeric_cols[i]) 
}

#Calculate ICC df
icc_df <- data.frame(icc = rep(NA, length(numeric_cols)), 
                     eta_sq = rep(NA, length(numeric_cols)),
                     sem = rep(NA, length(numeric_cols)))

icc_df$dv = row.names(icc_df)
row.names(icc_df) = seq(1:nrow(icc_df))
icc_df$task = 'task'
icc_df[grep('survey', icc_df$dv), 'task'] = 'survey'
icc_df = icc_df %>%
  select(dv, task, icc, eta_sq, sem)
row.names(icc_df) = NULL

@@

Average reliability metrics and the number of variables for both the surveys and the tasks
Both Spearman $\rho$'s as well the ICC's are higher for surveys.  
Systematic differences between the two measurements ($\eta^2$) are higher for tasks than surveys (potentially indicating learning effects?).  
Unexpectedly, median standard error of measurement is higher for surveys but the distribution below suggests that this is a side effect of having too many task measures with very low SEMs.
```{r}
datatable(rel_df %>%
  group_by(task) %>%
  summarise(median_spearman = median(spearman),
            median_icc = median(icc),
            median_eta_sq = median(eta_sq),
            median_sem = median(sem),
            num_vars = n()))
```

*Rank orders for tasks*

```{r}

*Rank order for surveys:*
- Demographics are excluded from this list of variables (because they are saved for predictive analyses).  
- Holt and Laury test-retest is the worst for *surveys*. But note that this is not a survey that has been studied extensively psychometrically like others. This is more similar to our tasks (one that is used more often by economists). In previous data I have worked with retest reliablities for this task was ~.35 so the results here are in line with this.  
- Otherwise the lowest survey reliabilities are 0.5.
```{r}

### Histograms of reliability metrics

@@

#### Spearman

@@

### ICC

@@

### Partial $\eta^2$

@@

### SEM

@@

bootstrap_relialibility = function(metric = c('icc', 'spearman', 'eta_sq', 'sem'), dv_var){
  tmp_sample = sample_workers()
  get_icc(dv_var, sample = 'bootstrap', sample_vec = tmp_sample, format='wide')
  get_spearman(dv_var, sample = 'bootstrap', sample_vec = tmp_sample, format='wide')
  get_eta(dv_var, sample = 'bootstrap', sample_vec = tmp_sample)
  get_sem(dv_var, sample = 'bootstrap', sample_vec = tmp_sample, format='wide')
}

sample_workers()
